Ab Initio Effective One-Electron Potentials for Repulsive and Charge-Transfer Interaction Energies at Hartree-Fock Level

### 2019-05-20 papers/2_oep_technique/paper-r0.tex
Mistakes in documentation corrected. Paper II adds

\section{\label{s:1}Introduction}
\section{\label{s:2}Theory}

Solution of the Schr{\"o}dinger equation provides the wavefunction
of the molecular aggregate $AB$,
%
\begin{equation}
 E^{AB} = \tBraKet{\Psi^{AB}}{\mathscr{H}}{\Psi^{AB}} \;,
\end{equation}
%
from which any property $P$, understood as an expectation value of the
associated operator $\mathscr{P}$, can be computed according to
%
\begin{equation}
 P = \tBraKet{\Psi^{AB}}{\mathscr{P}}{\Psi^{AB}} \;.
\end{equation}
%

\subsection{\label{ss:2.1}Intermolecular Interaction-Induced Observable}

Within a supermolecular approch, the part of a property that arises due to the
intermolecular interaction between $A$ and $B$ is given by
%
\begin{equation} \label{e:delta-p}
	\Delta^{AB} P = \tBraKet{\Psi^{AB}}{\mathscr{P}}{\Psi^{AB}} - 
	    \tBraKet{\Psi^{A}}{\mathscr{P}}{\Psi^{A}} -
	    \tBraKet{\Psi^{B}}{\mathscr{P}}{\Psi^{B}}\;.
\end{equation}
%
In a certain basis, expectation values of any operator $\mathscr{O}$ evaluated for
any system $X$ can be formally
expanded in a following series
%
\begin{equation} \label{e:exp-val-series}
	\tBraKet{\Psi^{X}}{\mathscr{O}}{\Psi^{X}} =
	\sum_{ij} C^{X}_{ij} \tBraKet{i}{\hat{h}}{j}
	+ \sum_{ijkl} C^{X}_{ijkl} \BraKet{ij}{kl}
\end{equation}
%
where $\hat{h}$ is the core Hamiltonian one\hyp{}electron operator
and the tensors
${\bf C}^{(2)}$ and ${\bf C}^{(4)}$ are characteristic for the property of interest.
Consider an approximation in which the basis functions $\phi_i$ 
are localized on either molecule. 
For this case, application of Eq.~\eqref{e:exp-val-series}
to Eq.~\eqref{e:delta-p} yields the partitioning of the interaction\hyp{}induced property,
for which the one\hyp{}electron part reads
%
\begin{equation} \label{e:delta-p-part-1el}
	\sum_{ij} \Big\{ 
	C^{AA}_{ij}  \BraKet{A}{A} + %\tBraKet{i^A}{\hat{h}}{j^A} + 
	C^{BB}_{ij}  \BraKet{B}{B} + %\tBraKet{i^B}{\hat{h}}{j^B} + 2 
2	C^{AB}_{ij}  \BraKet{A}{B}   %\tBraKet{i^A}{\hat{h}}{j^B} 
	\Big\}
\end{equation}
%
and the two\hyp{}electron part reads
%
\begin{multline} \label{e:delta-p-part-2el}
	\sum_{ijkl} \Big\{  
	C^{AAAA}_{ijkl} \BraKet{AA}{AA} +    %\BraKet{i^Aj^A}{k^Al^A} +
	C^{BBBB}_{ijkl} \BraKet{BB}{BB} + \\ %\BraKet{i^Bj^B}{k^Bl^B} +
	{s} \left[ C^{AABB}_{ijkl} \BraKet{AA}{BB} \right] + 
	{s} \left[ C^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] + \\
	{s} \left[ C^{AAAB}_{ijkl} \BraKet{AA}{AB} \right] +
	{s} \left[ C^{ABBB}_{ijkl} \BraKet{AB}{BB} \right]
	\Big\}
\end{multline}
%
To simplify the resulting expressions
in Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el},
the shorthand notation $\BraKet{A}{B} \equiv \tBraKet{i^A}{\hat{h}}{j^B}$ and 
$\BraKet{AB}{CD} \equiv \BraKet{i^Aj^B}{k^Cl^D}$
was introduced, and
the real orbitals 
were assumed to simplify the symmetry of the one\hyp{} and two\hyp{}electron integrals.
The operator $s$ permutes and sums over all the combinations of the basis function indices
given their partitioning scheme between molecules $A$ and $B$.

>>>>

\section{\label{s:1}Introduction}
\section{\label{s:2}Effective One-Electron Potentials: Configuration Interaction}

The time\hyp{}independent Schr{\"o}dinger equation provides the stationary solution to the
electronic state
of a molecular system in terms of its $N$\hyp{}electron wavefunction, $\Psi$,
%
\begin{equation} \label{e:schrodinger}
 E = \tBraKet{\Psi}{\mathscr{H}}{\Psi} \;,
\end{equation}
%
for a quantum Hamiltonian, $\mathscr{H} = \hat{h}(1) + \hat{o}(2)$,
with $\hat{h}(1)$ being the core Hamiltonian one\hyp{}electron operator
and $\hat{o}(2)$ the two\hyp{}electron repulsion operator.
From the above solution, any property $P$, understood as an expectation value of the
associated operator $\mathscr{P}$, can be computed according to
%
\begin{equation} \label{e:prop}
 P = \tBraKet{\Psi}{\mathscr{P}}{\Psi} \;.
\end{equation}
%
The exact solution to Eq.~\eqref{e:schrodinger} is formally given by
the configuration interaction (CI) expansion around the reference 
(approximate) wavefunction, $\Psi^{(0)}$,
typically the solution to the Hartree\hyp{}Fock (HF) problem,
%
\begin{equation} \label{e:ci}
 \Ket{\Psi} = \Ket{\Psi^{(0)}} + \sum_{ra} C_{r}^{a} \Ket{\Psi_{r}^{a}} + 
	 \sum_{rsab} C_{rs}^{ab} \Ket{\Psi_{rs}^{ab}} + \ldots
\end{equation}
%
In the above equation,
$\Ket{\Psi_{rs\ldots}^{ab\ldots}}$ are called the CI configurations
with the associated CI coefficients $C_{rs\ldots}^{ab\ldots}$, created by `exciting'
one or more electrons from the $a$th ($b$th) occupied molecular orbital to the $a$th ($b$th)
virtual (unoccupied) molecular orbital, obtained from the HF solution.
By inserting Eq.~\eqref{e:ci} into Eq.~\eqref{e:prop} 
and using the Slater\hyp{}Condon rules for evaluating the one\hyp{} and two\hyp{}electron
operator matrix element between
the CI configurations, it can be shown that the following holds
%
\begin{equation} \label{e:exp-val-series}
	P =
	\sum_{ij} P_{ij} \tBraKet{i}{\hat{h}}{j}
	+ \sum_{ijkl} P_{ijkl} \BraKet{ij}{kl}  \;,
\end{equation}
%
where $\tBraKet{i}{\hat{h}}{j} \equiv \int \phi^*_i({\bf r}) \hat{h} \phi_j({\bf r}) d{\bf r} $,
$\BraKet{ij}{kl}$ is the electron\hyp{}repulsion integral (ERI) defined by
%
\begin{equation}
	\BraKet{ij}{kl} \equiv
	\iint 
	\frac{ \phi_i^{*}({\bf r}_1) \phi_j({\bf r}_1) 
	       \phi_k^{*}({\bf r}_2) \phi_l({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2
\end{equation}
%
whereas the second\hyp{}rank tensor ${\bf P}^{(2)}$ 
and the fourth\hyp{}rank tensor ${\bf P}^{(4)}$ 
are well defined and characteristic for the property of interest.
%In this work we are interested in cases when

%$\Psi^{AB}$. In this Section we consider theories to calculate the effect of the interaction
%of molecules $A$ and $B$ on the property $P$ knowing only the unperturbed wavefunctions,
%$\Psi^{A}$ and $\Psi^{B}$.

\subsection{\label{ss:2.1}Intermolecular Interaction-Induced Property: Localized Basis Set Approximation}

Within a supermolecular approch, the part of a property that arises due to the
intermolecular interaction between $A$ and $B$ is given by
%
\begin{equation} \label{e:delta-p}
	\Delta^{AB} P = \tBraKet{\Psi^{AB}}{\mathscr{P}}{\Psi^{AB}} - 
	    \tBraKet{\Psi^{A}}{\mathscr{P}}{\Psi^{A}} -
	    \tBraKet{\Psi^{B}}{\mathscr{P}}{\Psi^{B}}\;.
\end{equation}
%
where $\Psi^{AB}$ and $\Psi^{A(B)}$ are the wavefunctions of the interacting molecular aggregate
and the unperturbed (non\hyp{}interacting) molecules, respectively.

Consider an approximation in which the basis functions $\phi_i$ 
are localized on either molecule. 
For this case, application of Eq.~\eqref{e:exp-val-series}
to Eq.~\eqref{e:delta-p} yields the partitioning of the interaction\hyp{}induced property,
for which the one\hyp{}electron part reads
%
\begin{equation} \label{e:delta-p-part-1el}
	\sum_{ij} \Big\{ 
	C^{AA}_{ij}  \BraKet{A}{A} + %\tBraKet{i^A}{\hat{h}}{j^A} + 
	C^{BB}_{ij}  \BraKet{B}{B} + %\tBraKet{i^B}{\hat{h}}{j^B} + 2 
2	C^{AB}_{ij}  \BraKet{A}{B}   %\tBraKet{i^A}{\hat{h}}{j^B} 
	\Big\}
\end{equation}
%
and the two\hyp{}electron part reads
%
\begin{multline} \label{e:delta-p-part-2el}
	\sum_{ijkl} \Big\{  
	C^{AAAA}_{ijkl} \BraKet{AA}{AA} +    %\BraKet{i^Aj^A}{k^Al^A} +
	C^{BBBB}_{ijkl} \BraKet{BB}{BB} + \\ %\BraKet{i^Bj^B}{k^Bl^B} +
	{s} \left[ C^{AABB}_{ijkl} \BraKet{AA}{BB} \right] + 
	{s} \left[ C^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] + \\
	{s} \left[ C^{AAAB}_{ijkl} \BraKet{AA}{AB} \right] +
	{s} \left[ C^{ABBB}_{ijkl} \BraKet{AB}{BB} \right]
	\Big\}
\end{multline}
%
To simplify the resulting expressions
in Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el},
the shorthand notation $\BraKet{A}{B} \equiv \tBraKet{i^A}{\hat{h}}{j^B}$ and 
$\BraKet{AB}{CD} \equiv \BraKet{i^Aj^B}{k^Cl^D}$
was introduced, and
the real orbitals 
were assumed to simplify the symmetry of the one\hyp{} and two\hyp{}electron integrals.
The operator $s$ permutes and sums over all the combinations of the basis function indices
given their partitioning scheme between molecules $A$ and $B$.
